---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
##Libraries used
```{r}
library(ggplot2)
library(dplyr)
library(gridExtra)
Sys.setlocale("LC_TIME", "C")

```

## Loading and preprocessing the data
1.Load the data (i.e. \color{red}{\verb|read.csv()|}read.csv())
```{r}
data <- read.table(unzip("activity.zip","activity.csv"),header=T, sep=",", stringsAsFactors = FALSE)
```

2.Process/transform the data (if necessary) into a format suitable for your analysis
```{r}
data$date <- as.Date(data$date, format= "%Y-%m-%d")
```


## What is mean total number of steps taken per day?
1. Calculate the total number of steps taken per day
```{r}
stepsperday <- aggregate(steps ~ date,data , sum)
print(stepsperday)
```

2. If you do not understand the difference between a histogram and a barplot, research the difference between them. Make a histogram of the total number of steps taken each day
```{r}
g1 <- ggplot(stepsperday, aes(steps, fill=..count..)) + geom_histogram(bins=40) + scale_fill_gradient("Count", low="green", high=  "blue") + ggtitle("Histogram total steps taken per day")
g1
```

3.Calculate and report the mean and median of the total number of steps taken per day
```{r}
mediansteps <- median(stepsperday$steps)
meansteps <- mean(stepsperday$steps)
print(paste("Mean of the total number of steps taken per day:",mediansteps))
print(paste("Median of the total number of steps taken per day:",as.integer(meansteps)))
```

## What is the average daily activity pattern?
Make a time series plot (i.e. \color{red}{\verb|type = "l"|}type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
stepsperinterval <- aggregate(steps ~ interval,data , mean)
ggplot(stepsperinterval, aes(x=interval, y=steps))+ geom_line(col="orange")+xlab("Time (Min)")+ylab("Steps average per day")+ ggtitle("AVG. NO. of steps per time interval")
```

2.Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?
```{r}
print(stepsperinterval[stepsperinterval$steps == max(stepsperinterval$steps),])
```

## Imputing missing values
1.Calculate and report the total number of missing values in the dataset (i.e. the total number of rows with NAs)
```{r}
nas <- sum(is.na(data$steps))
print(paste("Total missing values from steps:", nas))
```

2.Devise a strategy for filling in all of the missing values in the dataset. The strategy does not need to be sophisticated. For example, you could use the mean/median for that day, or the mean for that 5-minute interval, etc.
3.Create a new dataset that is equal to the original dataset but with the missing data filled in.
```{r}
nonastepsperday <- data %>% replace(is.na(.),mean(data$steps,na.rm=T))
nonastepsperday <- aggregate(steps ~ date,nonastepsperday , sum)
```
4.Make a histogram of the total number of steps taken each day 
```{r}
g2 <- ggplot(nonastepsperday, aes(steps, fill=..count..)) + geom_histogram(bins=40) + scale_fill_gradient("Count", low="green", high=  "blue") + ggtitle("Histogram total steps taken per day NA replaced")
g2
```
and Calculate and report the mean and median total number of steps taken per day. 
```{r}
mediansteps2 <- median(nonastepsperday$steps)
meansteps2 <- mean(nonastepsperday$steps)
print(paste("Mean of the total number of steps taken per day:",as.integer(mediansteps2)))
print(paste("Median of the total number of steps taken per day:",as.integer(meansteps2)))
```
Do these values differ from the estimates from the first part of the assignment? 
```{r}
print(paste("Median differ?",mediansteps2 != mediansteps))
print(paste("Mean differ?",meansteps2  != meansteps))


```

What is the impact of imputing missing data on the estimates of the total daily number of steps?
```{r}
grid.arrange(g1,g2, nrow=1)
```



```{r}
nonastepsperday2 <- nonastepsperday %>% mutate(fill = "NoNas")
stepsperday2 <- stepsperday %>%  mutate(fill= "Nas")
nasnonas <- rbind(nonastepsperday2, stepsperday2)
ggplot(nasnonas, aes(steps)) + geom_histogram(data=subset(nasnonas, fill=="NoNas"),bins=40, fill="blue", alpha=0.3)  + geom_histogram(data=subset(nasnonas,fill == "Nas"),bins=40, fill="red", alpha=0.3) +
        ggtitle("Superposition of total steps taken per day histograms")
```
As we can see, there is some impact.


## Are there differences in activity patterns between weekdays and weekends?

1.Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.
```{r}
data2 <- data %>% replace(is.na(.),mean(data$steps,na.rm=T))
data2 <- data2 %>% mutate(weekday= weekdays(date))
data2$weekday[!(weekdays(as.Date(data2$date)) %in% c('Saturday','Sunday'))] <- "weekday"
data2$weekday[(weekdays(as.Date(data2$date)) %in% c('Saturday','Sunday'))] <- "weekend"

```

2.Make a panel plot containing a time series plot (i.e. \color{red}{\verb|type = "l"|}type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all weekday days or weekend days (y-axis). See the README file in the GitHub repository to see an example of what this plot should look like using simulated data.
```{r}
dataweekdays <- aggregate(steps ~ interval ,data2[data2$weekday=="weekday",] , mean)
dataweekdays <- mutate(dataweekdays,weekday="weekday")
dataweekends <- aggregate(steps ~ interval ,data2[data2$weekday=="weekend",] , mean)
dataweekends <- mutate(dataweekends, weekday="weekend")
rbdata <- rbind(dataweekends, dataweekdays)
ggplot(rbdata, aes(x=interval, y=steps))+ geom_line(col="orange")+xlab("Interval")+
        ylab("Number of steps")+
        facet_wrap(~weekday, dir="v") 

```


